---
title: "HW 02"
author: "INSERT YOUR NAME HERE"
format: 
  html:
    embed-resources: true
toc: true
---

```{r}
#| label: load-pkgs
#| message: false
#| echo: false

if (!require("pacman")) 
  install.packages("pacman, ggridges")

pacman::p_load(tidyverse,
               ggridges,
               viridis,
               dsbox)

```

```{r}
edibnb |>
  glimpse()
```

## 1 - A new day, a new plot, a new geom

```{r}

# filter and reorder
edibnb_clean <- dsbox::edibnb |>
  filter(!is.na(review_scores_rating), !is.na(neighbourhood)) |>
  mutate(
    neighbourhood_ordered = fct_reorder(
      neighbourhood,
      review_scores_rating,
      .fun  = median
    )
  )

# plot
ggplot(edibnb_clean, aes(
    x = review_scores_rating,
    y = neighbourhood_ordered
  )) +
  geom_density_ridges(
    fill  = "steelblue",
    color = "white",
    alpha = 0.7,
    scale = 1.3
  ) +
  labs(
    title   = "Airbnb Review Scores by Edinburgh Neighborhood",
    x       = "Review Score",
    y       = "Neighborhood (highest median at top)",
    caption = "Source: edibnb from the dsbox library"
  ) +
  theme_ridges(grid = TRUE) +
  theme(
    plot.title      = element_text(hjust = 0),  # left align
    legend.position = "none"
  )



```

Most neighborhoods examined show generally high reviews. The neighborhoods with the highest medians display tighter clusters communicating consistently high guest satisfaction. The lower scored neighborhoods have flatter more spread plots meaning a wider variety of satisfaction. Outliers on the low end of the scale force a lot of white space.

## 2 - Foreign Connected PACs

```{r}
#| label: label-me-2

```

## 3 - Median housing prices in the US

## 4 - Expect More. Plot More.

## 5 - Mirror, mirror on the wall, who’s the ugliest of them all?
